---
title: "Assignment 1"
author: "Test Student"
date: "2025-06-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal of this assignment  

The goal of this assignment is to (i) review a few basic concepts and to (ii) make sure that you can knit an `Rmd` document with the basic features that will be needed throughout this course. 
Please complete exercises (1) and (2), rename the `Rmd` file to "Assignment1_YourLastName.Rmd" (e.g., "Assignment1_Smith.Rmd"), your name in the header, and knit the Rmd to an html file or pdf file. 
Please submit that html or pdf file on CANVAS by Friday, June 20th midnight. 
You may work in pairs, but each one of you will have to submit your own file. 

## 1. Using your own words, describe a randomized complete block design (RCBD) in 3-5 sentences. Then write out a statistical model (using mathematical notation) that could be fitted to data generated by an RCBD.   

An RCBD is a type of design that gathers groups of similar EUs and accounts for their variability....

A statistical model that could be connected to an RCBD is 

$$y_{ijk} = \mu + T_j + B_k + \varepsilon_{ijk}, \varepsilon_{ijk}\sim N(0, \sigma^2),\ $$

where:
- $\mu$ is the overall mean, 
- $T_j$ is the effect of the $j$th treatment, 
- $B_k$ is the effect of the $k$th block, 
- $\varepsilon_{ijk}$ is the residual of the $i$th observation, $j$th treatment, $k$th block.

## 2. Edit the R code below:  

- Silence the `messages` and `warnings`: your submitted pdf or html should include the code, but not the warnings/messages seen below. 
- Print a summary of `model1`. 
- Find and print the value for $\hat\sigma^2$ estimated in `model1`. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(agridat)
data("omer.sorghum")
df <- omer.sorghum %>% filter(env == "E3")
model1 <- lm(yield ~ 1 + gen + rep, data = df)
```

```{r}
summary(model1)
```

```{r}
sigma(model1)^2
sum(model1$residuals^2) / model1$df.residual
summary(model1)$sigma^2
```

