---
title: "Effect of Temperature and Banana quantity on Muffin Height"
author: "J Lacasa"
date: "2025-06-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background  

Banana muffins, much like banana bread, present a unique baking challenge due to the inclusion of mashed bananas in the batter. While this ingredient contributes moisture and flavor, it can also result in a dense crumb if not balanced properly with other recipe components and baking conditions. In particular, oven temperature and banana quantity play a critical role in achieving a light, fluffy texture. This study begins with a standard banana muffin recipe and explores how adjusting the baking temperature affects the final crumb structure, with the goal of identifying optimal conditions for producing well-risen, tender muffins.
**The objective of this study is to find the best combination of oven temperature and banana quantity to bake muffins.** 

## Materials  

The facility that is available to bake the muffins includes.

## ANOVA table 

```{r echo=FALSE}
t_design <- data.frame(Source = c("Day (block)", "", "Error(oven)", "Total"),
                       df = c("r-1 = 3-1 = 2", " ", "(t-1)r = (3-1)*3 = 6", "N-1 = 8"))

t_trt <- data.frame(Source = c(" ", "Temperature", "Parallels", "Total"),
                    df = c(" ", "t-1 = 3-1 = 2", "N-t = 9-3 = 6", "N-1 = 8"))

t_rcbd <- data.frame(Source = c("Day", "Temperature", "Error(oven)", "Total"),
                    df = c("r-1 = 3-1 = 2", "t-1 = 3-1 = 2", "(t-1)r - (t-1) = 2*3 - 2 = 4", "N-1 = 8"))

knitr::kable(t_design, caption = "Design or Topographical Sources of Variability")
knitr::kable(t_trt, caption = "Treatment Sources of Variability")
knitr::kable(t_rcbd, caption = "Combined Table of the Sources of Variability")
```

## References  

